---
title: "Bio302 Practical 1 Exploratory Data Analysis"
author: "Richard J. Telford"
date: "May 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1 Peguins

Load the `palmerpenguins` package and see the `penguins` data.

Loading packages and viewing data
```{r}
library(tidyverse)
library(palmerpenguins)
view(penguins)
```

##Code from Richard (next day review) 

cor.test(~ bill_length_mm + bill_depth_mm, data = penguins, use = "dataframe",
subset = speciess == "Gentoo")

penguins %>% 
group_by(species) %>% 
summarise(cor = cor( bill_length_mm, bill_depth_mm, use="dataframe"))


broom::glance pulls out results and makes it into a tibble. If you do a correlation test and get results, you can use broom::glance. 

Find the range, mean, median and variance of variable `bill_length_mm`.

##Finding the mean
```{r}
mean_peng <-penguins %>% 
  select(bill_length_mm) %>%
  summarise(mean = mean(bill_length_mm, na.rm = TRUE)) 
  head(mean_peng)
```

##Finding the median
```{r}
 median_peng <- penguins %>% 
  select(bill_length_mm) %>%
  summarise(median = median(bill_length_mm, na.rm = TRUE))
  head(median_peng)



```

##Finding the variance
```{r}
 var_peng <- penguins %>% 
  select(bill_length_mm) %>%
  summarise(variance = var(bill_length_mm, na.rm = TRUE))
  
head(var_peng)





```


What is the standard error of the mean of `bill_length_mm`.

Finding the SD first, then the standard error
```{r}
#Selecting only the column that needs to be counted
bill_peng <- penguins %>% 
  select(bill_length_mm) %>% 
  na.exclude()

#Counting number of observations for bill_Length_mm
#tally(p2)

#SE = SD/sqrt(n)

#SD = sqrt(var(y))
#sqrt(p3)

standard_error <- sqrt(var_peng)/sqrt(tally(bill_peng))
head(standard_error)
#Standard error is 0.294
```


Plot a histogram of variable `bill_length_mm`.

Plotting histogram and marking the position of the mean and median.
```{r}
ggplot(bill_peng, aes(x = bill_length_mm)) +
  geom_histogram() +
  geom_vline(xintercept = mean(penguins$bill_length_mm, na.rm = TRUE), colour = "red") +
geom_vline(xintercept = median(penguins$bill_length_mm, na.rm = TRUE), colour = "blue")
```

Mark the position of the mean and median.

Plot `bill_length_mm` against `bill_depth_mm`

Find the covariance and correlation of `bill_length_mm` against `bill_depth_mm`.
```{r}
#Finding correlation between bill length and bill depth.
penguins %>% 
group_by(species) %>% 
na.exclude() %>% 
summarise(cor = cor( bill_length_mm, bill_depth_mm)) 
  
#Finding covariance between bill length and bill depth.
penguins %>% 
group_by(species) %>% 
na.exclude() %>% 
summarise(cov = cov( bill_length_mm, bill_depth_mm)) 

#penguins %>% 
#  group_by(species) %>% 
#  summarise(cor = broom::glance(cor.test(bill_length_mm, #bill_depth_mm)))

```


Is the correlation significant? (hint use cor.test())

How should the p-value be interpreted?

There are three levels in factor `species` Plot `bill_length_mm` against `bill_depth_mm`, coloured by `species`

Is the correlation between `bill_length_mm` against `bill_depth_mm` significant for any of the groups?

This is a sub-group analysis, what are the dangers of this type of analysis?

# Part 2 not penguins

Install the `datasauRus` package with `install.packages("datasauRus")` and load it with `library`.

Dataset `datasaurus_dozen` has `x` and `y` variables and a dataset name. 
For each dataset, calculate the mean and standard deviation of x and y and the correlationn between x and y.