---
title: "Bio302 Practical 1 Exploratory Data Analysis"
author: "Richard J. Telford"
date: "May 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1 Peguins

Load the `palmerpenguins` package and see the `penguins` data.

Loading packages and viewing data
```{r}
library(tidyverse)
library(palmerpenguins)
view(penguins)
```


Find the range, mean, median and variance of variable `bill_length_mm`.

##Finding the mean
```{r}
penguins %>% 
  select(bill_length_mm) %>%
  summarise(mean = mean(bill_length_mm, na.rm = TRUE)) %>% 
  head()
```

##Finding the median
```{r}
 penguins %>% 
  select(bill_length_mm) %>%
  summarise(median = median(bill_length_mm, na.rm = TRUE)) %>% 
  head()



```

##Finding the variance
```{r}
 p3 <- penguins %>% 
  select(bill_length_mm) %>%
  summarise(variance = var(bill_length_mm, na.rm = TRUE)) 

  head(p3)




```


What is the standard error of the mean of `bill_length_mm`.

Finding the SD first, then the standard error
```{r}
#Selecting only the column that needs to be counted
p2 <- penguins %>% 
  select(bill_length_mm)

#Counting number of observations for bill_Length_mm
#tally(p2)

#SE = SD/sqrt(n)

#SD = sqrt(var(y))
#sqrt(p3)

standard_error <- sqrt(p3)/sqrt(tally(p2))
head(standard_error)
#Standard error is 0.294
```


Plot a histogram of variable `bill_length_mm`.
```{r}
ggplot(p2, aes(x = bill_length_mm, y = biomass, fill = site)) +
  geom_histogram()
```

Mark the position of the mean and median.

Plot `bill_length_mm` against `bill_depth_mm`

Find the covariance and correlation of `bill_length_mm` against `bill_depth_mm`.

Is the correlation significant? (hint use cor.test())

How should the p-value be interpreted?

There are three levels in factor `species` Plot `bill_length_mm` against `bill_depth_mm`, coloured by `species`

Is the correlation between `bill_length_mm` against `bill_depth_mm` significant for any of the groups?

This is a sub-group analysis, what are the dangers of this type of analysis?

# Part 2 not penguins

Install the `datasauRus` package with `install.packages("datasauRus")` and load it with `library`.

Dataset `datasaurus_dozen` has `x` and `y` variables and a dataset name. 
For each dataset, calculate the mean and standard deviation of x and y and the correlationn between x and y.