---
title: "Linear models"
author: "Richard J. Telford"
date: "May 26, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear Modelling Exercises

##1. Import the penguin data from the `palmerpenguin` package

##Loading packages

```{r}
library(qpcR)
library(tidyverse)
library(palmerpenguins)
head(penguins)

```

##2. Do big birds have big beaks? Fit a linear model between bill length and body mass, and interpret the coefficients.
```{r}
 length_bodymass.df <- penguins %>% 
  select(bill_length_mm, body_mass_g) %>% 
  na.exclude()
head (length_bodymass.df)

fit.lm <- lm(bill_length_mm~body_mass_g, data = length_bodymass.df)

#plot(length_bodymass.df$bill_length_mm, length_bodymass.df$body_mass_g)
#lines(length_bodymass.df$bill_length_mm, fitted(fit.lm), col = "red")

summary(fit.lm)

ggplot(length_bodymass.df) +
  aes(x = bill_length_mm, y = body_mass_g) +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  geom_smooth(method = lm,formula = y~x, colour = "red")
  theme_minimal()
  
par(mfrow=c(2,2))  
plot(fit.lm)
  
#Residual vs Fitted shows a flat line (good)
#Normal Q-Q seems to follow line?
#Scale_location: somewhat flat line


```


##3. Examine the diagnostics plots. Is everything OK?

Yes

##3. Calculate the residual sum of squares (find sum the squared difference between the estimates calculated from the coefficients and the observed values).

```{r}
deviance(fit.lm)

#Package that is supposed to calculate RSS
#RSS(fit.lm)


```



4. Recalculate the Residual sum of squares using different values for the coefficients (in the formula yi = Beta + Beta xi + εi).

```{r}

```


5. Is the new RSS larger or smaller?
6. Estimate the slope using covariance and variance.
7. Do you get the same result?
##8. Add species to the model as a predictor + 9. Does species improve the model?

```{r}
fit1.lm <- lm(bill_length_mm~body_mass_g + species, data = penguins)

summary(fit1.lm)

anova(fit.lm, fit1.lm)


```
 ***Model 2 has a lower RSS meaning it is bettet***

9. Does species improve the model?
10. How should the coefficients be interpreted?

