---
title: "Practical 3"
author: "Richard J. Telford"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## GLS

With the built-in `iris` data.

```{r}
library(tidyverse)
library(nlme)
```


Make a plot to show how `Petal.Length` varies between species.

```{r}
ggplot(iris, aes( x = Species, y = Sepal.Length)) +
  geom_boxplot()
```


Find the variance of `Petal.Length` for each species.

```{r}
iris %>% 
    group_by(Species) %>% 
  summarise(var = var(Sepal.Length))
```


Fit an anova using `lm` between `Petal.Length` and species and examine the diagnostic plots.

```{r}
iris.df <- lm(Petal.Length ~ Species, data = iris)
anova(iris.df)

par(mfrow=c(2,2))  
plot(iris.df)

```



Fit a `gls` for the same model. Have the coefficients changed? (remember to load ´nlme´ package)

```{r}
iris_gls.df <- gls(Petal.Length ~ Species, data = iris)
anova(iris_gls.df)

par(mfrow=c(2,2))  
plot(iris_gls.df)
```


Fit a `gls` for the same model but allow the variance to be different for each species by adding `varIdent` to the `weights` argument of the model. 

```{r}
#Why does this not work?
iris_gls1.df <- gls(Petal.Length ~ Species, data = iris, weights = varIdent(form = ~ 1|Species))

```


Use `AIC` to test if this is a better model.

```{r}


anova(iris_gls.df, iris_gls1.df)


```

