---
title: "Spatial and temporal structure"
author: "Richard J. Telford"
date: "June 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, message = FALSE)
```

1) Simulate a 100-observation autocorrelated timeseries with `arima.sim`, with a first order autoregressive coefficient of 0.5. Also make a time vector of `1:100`

```{r}
#making simulation


ts.sim <-arima.sim(list(order = c(1,0,0), #specifying that this is a AR1 model
                        ar = 0.5), #autocrolleation coeficient
                   n = 100) #number of observations

time.vector <- 1:100


```


2) Plot the data. 

```{r}
#plotting data
ts.plot(ts.sim) 
```


3) Regress the timeseries against time with an OLS model. Does the model appear to be statistically significant?

```{r}
#normally use a lm instead of gls
gls.1 <- gls(ts.sim ~ time.vector)

#anova(gls.1)

lm.1 <- lm(ts.sim ~time.vector)
anova(lm.1)
```


4) Plot the model diagnostics, including an `acf` and `pacf` of the residuals.

```{r}
#diagnostic plot
par(mfrow=c(2,2))
plot(lm.1)

#acf and pacf

acf(resid(lm.1))
pacf(resid(lm.1))


```


5) Use the Durbin-Watson test to test the residuals for autocorrelation.

```{r}
#Using Durbin-watson to test for autocorrelation

library(lmtest)
dwtest(lm.1)

#Showung positive autocorrelation for the residuals
```



6) Fit a gls with an appropriate correlation structure. Is this a better model? How have the p-value and effect size changed?

```{r}

gls.1 <- gls(ts.sim ~ time.vector)
summary(gls.1)

fit1.gls <- gls(ts.sim~time.vector, corr = corAR1())
summary(fit1.gls)

#anova(gls.1, fit1.gls)

```
The new model has a lower AIC, meaning it is better.


7) Repeat the above 1000 times and find how autocorrelation affects the distribution of the p-value and the effect size.

```{r}




library(broom.mixed)#package that allows glancing gls
re <-rerun(1000, {ts.sim <- arima.sim(list(order = c(1,0,0), 
                        ar = 0.5), 
                   n = 100)
time.vector <- 1:100
lm(ts.sim ~time.vector)
}) %>% 
  map_dfr(glance) # using glance to get usable information out

ggplot(data = re, aes(x = p.value))+ #choosing p-value to be shown in histogram
  geom_histogram()


```


## Real data

1) The built-in dataset LakeHuron has annual lake level data from 1885 to 1972
Load the data with the command data(LakeHuron)

```{r}
data("LakeHuron")
```


2) Plot the data.

```{r}
mod1 <- lm(LakeHuron~time(LakeHuron))
summary(mod1)

view(LakeHuron)
```


3) Regress the LakeHuron lake level against year using a linear model. Is there a significant trend?

```{r}
lm.2 <- lm(LakeHuron~time)
```


4) Plot the autocorrelation and partial autocorrelation functions for the residuals from this regression. Interpret them.

5) Fit an autoregressive models to the residuals. Compare the results with your interpretation of the PACF plot.


6) Fit a gls model using a corAR1 correlation structure. Test if the correlation structure is necessary. Is the trend significant? 

7) Fit a gls model using a corARMA correlation structure with two AR terms. Is this model an improvement?


